---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fsttable

<!-- badges: start -->
<!-- badges: end -->

R package `fsttable` aims to provide a fully functional `data.table` interface to on-disk `fst` files. The focus of the package is on keeping memory usage as low as possible woithout sacrificing features of in-memory `data.table` operations.

## Installation

You can install the latest package version with:

``` r
devtools::install_github("fstpackage/fsttable")
```

## Example

First, we create a on-disk _fst_ file containing a medium sized dataset:

```{r data}
library(fsttable)

# write some sample data to disk
nr_of_rows <- 1e6
x <- data.table::data.table(X = 1:nr_of_rows, Y = LETTERS[1 + (1:nr_of_rows) %% 26])
fst::write_fst(x, "1.fst")
```

Then we define our _fst\_table_ by using:

```{r proxy}
ft <- fst_table("1.fst")
```

This _fst\_table_ can be used as a regular _data.table_ object. For example, we can print:

```{r print}
ft
```

we can select columns:

```{r columns}
ft[, .(Y)]
```

and rows:

```{r rows}
ft[1:4,]
```

Or both at the same time:

```{r cols_and_rows}
ft[1:4, .(X)]
```

# Memory

During the operations shown above, the actual data was never fully loaded from the file. That's because of `fsttable`'s philosophy of keeping RAM usage as low as possible. Printing a few lines of a table doesn't require knowlegde of the remaining lines, so `fsttable` will never actualy load them.

Even when you create a new set:

```{r copy}
ft2 <- ft[1:4, .(X)]
```

No actual data is being loaded into RAM. The copy still uses the original _fst_ file to keep the data on-disk:

```{r internals}
# small size because actual data is still on disk
object.size(ft2)
```
